\bf Model Visualization Techniques for a Social Network Model

### 2017-03-14 paper.Rnw
most recent updates.

\section{Introduction} 
\st{
\begin{itemize}
\item What? - model visualization and SAOMs. briefly dsecribe what these things are
\item Why? - bring light to the RSiena blackbox. Need ways to visualize network models, collections of networks. 
\item How?
\end{itemize}
also talk about other dynamic network models briefly to put SAOMs in context.}

@@

\par The software SIENA, and it's \texttt{R} implementation \texttt{RSiena}, was developed by \citet{RSiena} to estimate the parameters of the SAOMs, among other social network models. This software is a huge contribution to the field of social network analysis, and when we began working with it we were eager to understand how it works. To better understand the \texttt{RSiena} software and the SAOMs themselves, we use the principals of model visualization from \citet{modelvis}. With the work of \citet{modelvis} guiding us, we explore RSiena and SAOMs with innovative visualizations, both static and dynamic. By bringing some light to the underlying methods and structures "behind the scenes" of SAOMs and RSiena, we will help other researchers working with the models better understand and better use these models.

\par The guiding principles of model visualization from \citet{modelvis} are (1) display the model in the data space, (2) collections of models are more informative than singletons, and (3) explore how algorithms work, instead of just looking at the final result. Stochastic actor-oriented models are a prime example of a set of models that  benefit greatly from application of model visualization. First, the models themselves include a continuous-time Markov chain (CTMC) that is completely hidden in the model fitting process. Bringing the CTMC out from behind-the-scenes and into the light of visualization  provides researchers with insights into the underlying features of the model. Secondly, SAOMs allow for a great deal of parameters that can be added to the objective function, each of which are attached to a network statistic. These statistics are often somewhat, if not highly, correlated, and so are the associated parameters. By visualizing collections of SAOMs, we can gain a better understanding of these correlations and find ways to deal with them and rectify their effects. Finally, estimation of the parameters in a SAOM relies on a Robbins-Monro algorithm, and convergence checks rely on simulations from the fitted model. Again, in these steps are largely hidden from the researcher computing them with software. We hope that by bringing these background elements and others into the foreground, stochastic actor-oriented models will be better understood and more accessible to social network analysts.

@@

\hh{mostly lit review plus a detailed description of the model and fitting process.}

>>>>

\section{Introduction} 
\st{
\begin{itemize}
%\item What? - model visualization and SAOMs. briefly dsecribe what these things are
%\item Why? - bring light to the RSiena blackbox. Need ways to visualize network models, collections of networks. 
\item How?
\end{itemize}
%also talk about other dynamic network models briefly to put SAOMs in context.
}

@@

\par The software SIENA, and its \texttt{R} implementation \texttt{RSiena}, was developed by \citet{RSiena} to estimate the parameters of the SAOMs, among other social network models. This software is a huge contribution to the field of social network analysis, and when we began working with it we were eager to understand how it works. To better understand the \texttt{RSiena} software and the SAOMs themselves, we use the principals of model visualization from \citet{modelvis}. With the work of \citet{modelvis} guiding us, we explore RSiena and SAOMs with innovative visualizations, both static and dynamic. By bringing some light to the underlying methods and structures "behind the scenes" of SAOMs and RSiena, we will help other researchers working with the models better understand and better use these models.

\par The guiding principles of model visualization from \citet{modelvis} are (1) display the model in the data space, (2) collections of models are more informative than singletons, and (3) explore how algorithms work, instead of just looking at the final result. Stochastic actor-oriented models are a prime example of a set of models that  benefit greatly from application of model visualization. First, the models themselves include a continuous-time Markov chain (CTMC) that is completely hidden in the model fitting process. Bringing the CTMC out from behind-the-scenes and into the light of visualization  provides researchers with insights into the underlying features of the model. Secondly, SAOMs allow for a great deal of parameters that can be added to the objective function, each of which are attached to a network statistic. These statistics are often somewhat, if not highly, correlated, and so are the associated parameters. By visualizing collections of SAOMs, we can gain a better understanding of these correlations and find ways to deal with them and rectify their effects. Finally, estimation of the parameters in a SAOM relies on a Robbins-Monro algorithm, and convergence checks rely on simulations from the fitted model. Again, in these steps are largely hidden from the researcher computing them with software. We hope that by bringing these background elements and others into the foregroundwith static and dynamic visualizations, stochastic actor-oriented models will be better understood and more accessible to social network analysts.

@@

%\hh{mostly lit review plus a detailed description of the model and fitting process.}
